Type Refinements for the Working Class

### 2015-11-26 paper.tex
rename & document macros

\newcommand\Refines[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset\OMode{#3}}
\newcommand\GRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}
\newcommand\RefinesCtx[2]{\IMode{#1}\sqsubset^\star\OMode{#2}}
\newcommand\EqRefines[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\sqsubset\OMode{#4}}

\newcommand\HRefines[4]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}\sqsubset\OMode{#4}}
\newcommand\HEqRefines[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\sqsubset\OMode{#5}}
\newcommand\PRefinesCtx[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset^\star\OMode{#3}}

\newcommand\RMemEq[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\in\IMode{#4}}
\newcommand\HRMemEq[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\in\IMode{#5}}

\newcommand\Pow[1]{\wp\left(#1\right)}

\newcommand\SubRefines[4]{\IMode{#1}\parallel\IMode{#2}\subseteq\IMode{#3}\sqsubset\OMode{#4}}
\newcommand\TyBase[1]{\mathtt{Base}_{#1}}
\newcommand\TyTop[1]{\top_{#1}}
\newcommand\TyBot[1]{\bot_{#1}}
\newcommand\TyMeet[1]{\cap_{#1}}
\newcommand\TyJoin[1]{\cup_{#1}}

@@

Our position is that these views of type theory are not in conflict, but rather
merely describe two distinct layers in a single, harmonious system. From this
perspective, formal type theories can do little more than negotiate matters of
grammar, and therefore may serve as a linguistic framework for mathematical
expression, being responsible for the management of variable binding and
substitution. On the other hand, mathematical objects find their \emph{meaning}
in the behavioral type theory, and are subject to an extensional equality.

@@

Then, we say that $\phi$ \emph{globally refines} $\tau$ (written
$\GRefines{\phi}{\tau}$) when we have $\Refines{\cdot}{\phi}{\tau}$.
Furthermore, when $\Refines{\Upsilon}{\phi}{\tau}$, for any renaming
\begin{tikzcd}[cramped]
  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd}
we can clearly define a new refinement $\Refines{\Upsilon'}{\phi\rho}{\tau}$.

@@

We will write $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ to mean that $\phi$ is
a \emph{subrefinement} of $\psi$.

@@

  To know $\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ (presupposing
  $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$) , is
  to know, for any renamings
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r, hook, "\IMode{\rho}"]
    & \IMode{\Upsilon'} \arrow[r, hook, "\IMode{\rho'}"]
    & \IMode{\Upsilon''}
  \end{tikzcd},
  that from $\RefEquate{\phi}{\rho}{M}{N}$ you can conclude
  $\RefEquate{\psi}{\rho'\circ\rho}{M\rho'}{N\rho'}$ for any
  $\Member{M,N}{\Exprs_\tau(\Upsilon')}$.

@@

Two refinements are equal when they denote the same PER. That is, we have
$\EqRefines{\Upsilon}{\phi}{\psi}{\tau}$ just when both
$\SubRefines{\Upsilon}{\phi}{\psi}{\tau}$ and
$\SubRefines{\Upsilon}{\psi}{\phi}{\tau}$.

@@

\newcommand\PRefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
\newcommand\PRefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
\newcommand\PFunRefNil{\vDash^\sqsubset_{\mathtt{nil}}}
\newcommand\PFunRefSnoc{\vDash^\sqsubset_{\mathtt{snoc}}}

Next, we define functional refinement
$\HEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
simultaneously with parametric context refinement
$\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and functional equality
$\HRMemEq{\Upsilon}{\Psi}{M}{N}{\phi}$. We will write
$\HRefines{\Upsilon}{\Psi}{\phi}{\tau}$ as a shorthand for
$\HEqRefines{\Upsilon}{\Psi}{\phi}{\phi}{\tau}$.

\newcommand\DefPRefinesCtxNil[1]{
  \infer[\PRefinesCtxNil]{
    \PRefinesCtx{#1}{\cdot}{\cdot}
  }{
  }
}
\newcommand\DefPRefinesCtxSnoc[6]{
  \infer[\PRefinesCtxSnoc]{
    \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
  }{
    \PRefinesCtx{#1}{#2}{#5} &
    \HRefines{#1}{#2}{#4}{#6}
  }
}

Parametric context refinement $\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ shall be
defined as follows, presupposing $\IsSymCtx{\Upsilon}$ and
$\IsVarCtx{\Gamma}$:
\begin{gather*}
  \DefPRefinesCtxNil{\Upsilon}\qquad
  \DefPRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\phi}{\Gamma}{\tau}
\end{gather*}

For a context refinement $\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and a renaming
\begin{tikzcd}[cramped]
  \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd},
we can define a new context refinement
$\PRefinesCtx{\Upsilon'}{\Psi\rho}{\Gamma}$ by applying $\rho$ pointwise at
each of the sort refinements in the context.

Functional refinement $\HEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ presupposes
$\PRefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
induction on the evidence for its first presupposition.

\paragraph{Case} $\DefPRefinesCtxNil{\Upsilon}$.
\[
  \infer[\PFunRefNil]{
    \HEqRefines{\Upsilon}{\cdot}{\phi}{\psi}{\tau}
  }{
    \EqRefines{\Upsilon}{\phi}{\psi}{\tau}
  }
\]

\paragraph{Case} $\DefPRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.

\begin{quote}
  To know $\HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}$ is to know, for any renaming
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}
  and closed terms $\Member{M,N}{\Exprs_\sigma(\Upsilon')}$, that from
  $\HRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}$, you can conclude
  $\HEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
  In other words:
  %
  \[
    \infer[\PFunRefSnoc]{
      \HEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
    }{
      \begin{array}{l}
        \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}.\
        \forall \Member{M,N}{\Exprs_\sigma(\Upsilon')}.\\
        \left(\HRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}\right)
        \Rightarrow
        \left(\HEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}\right)
      \end{array}
    }
  \]
\end{quote}

%%
%% \subsection{Functional Equality}
%%
%% Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
%% $\RefinesCtx{\Phi}{\Upsilon}$, $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$,
%% $\HRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
%% and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
%% induction on the evidence for the second presupposition
%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
%%
%% \paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
%% \[
%%   \infer{
%%     \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }
%% \]
%%
%% \paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
%% where $\RefinesCtx{\Phi}{\Upsilon}$.
%% \[
%%   \infer{
%%     \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
%%   }{
%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
%%       \HypJ{
%%         \HRMemEq{\Phi}{\Psi}{
%%           \Subst{\MV{m}}{x}{\phi}
%%         }{
%%           \Subst{\MV{n}}{x}{\psi}
%%         }{\phi}
%%       }{
%%         \HRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
%%       }
%%     }
%%   }
%% \]
%%
%% \section{Constructions on Refinements}
%%
%% Refinements can be arranged in a partial order via the
%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
%% at any sort $\tau$.
%%
%% We have the bottom refinement $\GRefines{\TyBot{\tau}}{\tau}$, which simply
%% contains no elements. Next, the top refinement $\GRefines{\TyTop{\tau}}{\tau}$
%% identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
%% and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
%% $\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
%% \begin{gather*}
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }\qquad
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \end{gather*}
%%
%% Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
%% dually:
%% \[
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi} &
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \]
%%
%% It is trivial to verify that these constructions on refinements give rise to a
%% lattice.
%%
%% \section{Multi-Sorted Computational Type Theory}
%%
%% Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
%% you specify how to form its equal canonical inhabitants, or values. Then, the
%% inhabitants of the types are explained uniformly by extending this relation
%% over an operational semantics such that two terms are equal members of a
%% type just when they compute to equal values of that type.
%%
%% More generally, one may wish to consider types which include computations in
%% addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
%% is a type, which would include all the members of $\psi$ as well as a divergent
%% computation. Another example is the type $\TyTop{\tau}$ which identifies all
%% closed terms of sort $\tau$.
%%
%% In order to ensure that such types are sensible, we must formulate a notion of
%% computational equivalence, and then say that a refinement is a \emph{type} just
%% when it respects computational equivalence; then, we are able to define types
%% as PERs on \emph{programs} in addition to values.

>>>>

% parametric refinement
\newcommand\Refines[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset\OMode{#3}}

% global parametric refinement
\newcommand\GloballyRefines[2]{\IMode{#1}\sqsubset\OMode{#2}}

% equality of parametric refinement
\newcommand\EqRefines[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\sqsubset\OMode{#4}}

% parametric context refinement
\newcommand\RefinesCtx[3]{\IMode{#1}\parallel\IMode{#2}\sqsubset^\star\OMode{#3}}

% functional refinement
\newcommand\FRefines[4]{%
  \IMode{#1}\parallel\IMode{#2}%
  \vDash\IMode{#3}\sqsubset\OMode{#4}%
}

% equality of functional refinement
\newcommand\FEqRefines[5]{%
  \IMode{#1}\parallel\IMode{#2}%
  \vDash\IMode{#3}=\IMode{#4}
  \sqsubset\OMode{#5}%
}

% parametric equality of refinement members
\newcommand\RMemEq[4]{\IMode{#1}\parallel\IMode{#2}=\IMode{#3}\in\IMode{#4}}

% functional equality of refinement members
\newcommand\FRMemEq[5]{\IMode{#1}\parallel\IMode{#2}\vDash\IMode{#3}=\IMode{#4}\in\IMode{#5}}

% parametric subrefinement
\newcommand\IsSubrefinement[4]{%
  \IMode{#1}\parallel\IMode{#2}%
  \subseteq\IMode{#3}\sqsubset\OMode{#4}%
}

% power object
\newcommand\Pow[1]{\wp\left(#1\right)}

\newcommand\TyBase[1]{\mathtt{Base}_{#1}}
\newcommand\TyTop[1]{\top_{#1}}
\newcommand\TyBot[1]{\bot_{#1}}
\newcommand\TyMeet[1]{\cap_{#1}}
\newcommand\TyJoin[1]{\cup_{#1}}

@@

Our position is that these views of type theory are not in conflict, but rather
describe two distinct layers in a single, harmonious system. From this
perspective, formal type theories can do little more than negotiate matters of
grammar, and therefore may serve as a linguistic framework for mathematical
expression, being responsible for the management of variable binding and
substitution. On the other hand, mathematical objects find their \emph{meaning}
in the behavioral type theory, and are subject to an extensional equality.

@@

Then, we say that $\phi$ \emph{globally refines} $\tau$ (written
$\GloballyRefines{\phi}{\tau}$) when we have $\Refines{\cdot}{\phi}{\tau}$.
Furthermore, when $\Refines{\Upsilon}{\phi}{\tau}$, for any renaming
\begin{tikzcd}[cramped]
  \IMode{\Upsilon}\arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd}
we can clearly define a new refinement $\Refines{\Upsilon'}{\phi\rho}{\tau}$.

@@

We will write $\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ to mean that $\phi$ is
a \emph{subrefinement} of $\psi$.

@@

  To know $\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ (presupposing
  $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$) , is
  to know, for any renamings
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r, hook, "\IMode{\rho}"]
    & \IMode{\Upsilon'} \arrow[r, hook, "\IMode{\rho'}"]
    & \IMode{\Upsilon''}
  \end{tikzcd},
  that from $\RefEquate{\phi}{\rho}{M}{N}$ you can conclude
  $\RefEquate{\psi}{\rho'\circ\rho}{M\rho'}{N\rho'}$ for any
  $\Member{M,N}{\Exprs_\tau(\Upsilon')}$.

@@

Two refinements are equal when they denote the same PER. That is, we have
$\EqRefines{\Upsilon}{\phi}{\psi}{\tau}$ just when both
$\IsSubrefinement{\Upsilon}{\phi}{\psi}{\tau}$ and
$\IsSubrefinement{\Upsilon}{\psi}{\phi}{\tau}$.

@@

\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
\newcommand\PFunRefNil{\vDash^\sqsubset_{\mathtt{nil}}}
\newcommand\PFunRefSnoc{\vDash^\sqsubset_{\mathtt{snoc}}}

Next, we define functional refinement
$\FEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
simultaneously with parametric context refinement
$\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and functional equality
$\FRMemEq{\Upsilon}{\Psi}{M}{N}{\phi}$. We will write
$\FRefines{\Upsilon}{\Psi}{\phi}{\tau}$ as a shorthand for
$\FEqRefines{\Upsilon}{\Psi}{\phi}{\phi}{\tau}$.

\newcommand\DefRefinesCtxNil[1]{
  \infer[\RefinesCtxNil]{
    \RefinesCtx{#1}{\cdot}{\cdot}
  }{
  }
}
\newcommand\DefRefinesCtxSnoc[6]{
  \infer[\RefinesCtxSnoc]{
    \RefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
  }{
    \RefinesCtx{#1}{#2}{#5} &
    \FRefines{#1}{#2}{#4}{#6}
  }
}

Parametric context refinement $\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ shall be
defined as follows, presupposing $\IsSymCtx{\Upsilon}$ and
$\IsVarCtx{\Gamma}$:
\begin{gather*}
  \DefRefinesCtxNil{\Upsilon}\qquad
  \DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\phi}{\Gamma}{\tau}
\end{gather*}

For a context refinement $\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and a renaming
\begin{tikzcd}[cramped]
  \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
\end{tikzcd},
we can define a new context refinement
$\RefinesCtx{\Upsilon'}{\Psi\rho}{\Gamma}$ by applying $\rho$ pointwise at
each of the sort refinements in the context.

Functional refinement $\FEqRefines{\Upsilon}{\Psi}{\phi}{\psi}{\tau}$ presupposes
$\RefinesCtx{\Upsilon}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
induction on the evidence for its first presupposition.

\paragraph{Case} $\DefRefinesCtxNil{\Upsilon}$.
\[
  \infer[\PFunRefNil]{
    \FEqRefines{\Upsilon}{\cdot}{\phi}{\psi}{\tau}
  }{
    \EqRefines{\Upsilon}{\phi}{\psi}{\tau}
  }
\]

\paragraph{Case} $\DefRefinesCtxSnoc{\Upsilon}{\Psi}{x}{\chi}{\Gamma}{\sigma}$.

\begin{quote}
  To know $\FEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}$ is to know, for any renaming
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}
  and closed terms $\Member{M,N}{\Exprs_\sigma(\Upsilon')}$, that from
  $\FRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}$, you can conclude
  $\FEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}$.
  In other words:
  %
  \[
    \infer[\PFunRefSnoc]{
      \FEqRefines{\Upsilon}{\Psi,x:\chi}{\phi}{\psi}{\tau}
    }{
      \begin{array}{l}
        \forall
        \begin{tikzcd}[cramped, ampersand replacement = \&]
          \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] \&\IMode{\Upsilon'}
        \end{tikzcd}.\
        \forall \Member{M,N}{\Exprs_\sigma(\Upsilon')}.\\
        \left(\FRMemEq{\Upsilon}{\Psi\rho}{M}{N}{\chi\rho}\right)
        \Rightarrow
        \left(\FEqRefines{\Upsilon}{\Psi\rho}{\Subst{M}{x}{\phi\rho}}{\Subst{N}{x}{\psi\rho}}{\tau}\right)
      \end{array}
    }
  \]
\end{quote}

%%
%% \subsection{Functional Equality}
%%
%% Functional equality $\FRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
%% $\RefinesCtx{\Phi}{\Upsilon}$, $\RefinesCtx{\Phi}{\Psi}{\Gamma}$,
%% $\FRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
%% and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
%% induction on the evidence for the second presupposition
%% $\RefinesCtx{\Phi}{\Psi}{\Gamma}$:
%%
%% \paragraph{Case} $\DefRefinesCtxNil{\Phi}$.
%% \[
%%   \infer{
%%     \FRMemEq{\Phi}{\cdot}{M}{N}{\phi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }
%% \]
%%
%% \paragraph{Case} $\DefRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
%% where $\RefinesCtx{\Phi}{\Upsilon}$.
%% \[
%%   \infer{
%%     \FRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
%%   }{
%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
%%       \HypJ{
%%         \FRMemEq{\Phi}{\Psi}{
%%           \Subst{\MV{m}}{x}{\phi}
%%         }{
%%           \Subst{\MV{n}}{x}{\psi}
%%         }{\phi}
%%       }{
%%         \FRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
%%       }
%%     }
%%   }
%% \]
%%
%% \section{Constructions on Refinements}
%%
%% Refinements can be arranged in a partial order via the
%% $\IsSubrefinement{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
%% at any sort $\tau$.
%%
%% We have the bottom refinement $\GloballyRefines{\TyBot{\tau}}{\tau}$, which simply
%% contains no elements. Next, the top refinement $\GloballyRefines{\TyTop{\tau}}{\tau}$
%% identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
%% and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
%% $\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
%% \begin{gather*}
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }\qquad
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \end{gather*}
%%
%% Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
%% dually:
%% \[
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi} &
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \]
%%
%% It is trivial to verify that these constructions on refinements give rise to a
%% lattice.
%%
%% \section{Multi-Sorted Computational Type Theory}
%%
%% Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
%% you specify how to form its equal canonical inhabitants, or values. Then, the
%% inhabitants of the types are explained uniformly by extending this relation
%% over an operational semantics such that two terms are equal members of a
%% type just when they compute to equal values of that type.
%%
%% More generally, one may wish to consider types which include computations in
%% addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
%% is a type, which would include all the members of $\psi$ as well as a divergent
%% computation. Another example is the type $\TyTop{\tau}$ which identifies all
%% closed terms of sort $\tau$.
%%
%% In order to ensure that such types are sensible, we must formulate a notion of
%% computational equivalence, and then say that a refinement is a \emph{type} just
%% when it respects computational equivalence; then, we are able to define types
%% as PERs on \emph{programs} in addition to values.
